Optimizing Abstract Abstract Machines

### 2012-11-14 paper/paper.tex

This paper is organized in two halfs: in the first, we start with a
quick review of the AAM approach to develop an analysis framework and
then apply our step-by-step optimization techniques in the simplified
setting of a core functional language.  This allows us to explicate
the optimizations with a minimal amount of inessential technical
overhead.  In the second half, we scale this approach up to an
analyzer for a realistic untyped, higher-order imperative language
with a number of interesting features.

At each step during the initial development, we evaluate the
implementation on a benchmark from Vardoulakis and
Shivers~\cite{dvanhorn:Vardoulakis2011CFA2} that tests distributivity
of multiplication over addition on Church numerals.
The example, while small, is informative:
\begin{enumerate}
\item it can be written in most modern programming languages,
%
\item it was designed to stress an analyzer's ability to deal with
  complicated environment and control structure arising from the use
  of higher-order functions to encode arithmetic, and
%
\item it proves to be the \emph{least} improved benchmark of the
  complete suite considered in section~\ref{sec:eval}, and thus it
  serves as a good sanity check and lower-bound for each of the
  optimization techniques considered.
\end{enumerate}

>>>>

This paper is organized in two halfs: in the first, we start with a
quick review of the AAM approach to develop an analysis framework and
then apply our step-by-step optimization techniques in the simplified
setting of a core functional language.  This allows us to explicate
the optimizations with a minimal amount of inessential technical
overhead.  In the second half, we scale this approach up to an
analyzer for a realistic untyped, higher-order imperative language
with a number of interesting features and then measure
improvements across a suite of benchmarks.

At each step during the initial presentation and development, we evaluate the
implementation on a benchmark from Vardoulakis and
Shivers~\cite{dvanhorn:Vardoulakis2011CFA2} that tests distributivity
of multiplication over addition on Church numerals.
For the step-by-step development, 
this benchmark is particularly informative:
\begin{enumerate}
\item it can be written in most modern programming languages,
%
\item it was designed to stress an analyzer's ability to deal with
  complicated environment and control structure arising from the use
  of higher-order functions to encode arithmetic, and
%
\item it proves to be the \emph{least} improved benchmark of the
  complete suite considered in section~\ref{sec:eval}, and thus it
  serves as a good sanity check and lower-bound for each of the
  optimization techniques considered.
\end{enumerate}
